\section{Recommendations for Further Research}
\label{sec:future_research}

Several topics could not be treated during this semester mostly because of the
time constraints.

\subsection{Shared Model Logic Code Injection}
\label{sub:model_code_injection}

One limitation of the \textit{Shared Model} is that the generated model classes
(POJOs) do not allow adding any logic to them. This would however be helpful in
order to share functionality in a way that is decoupled from any programming
language. One possibility to inject code into model objects is to use a high
level scripting language like Java Script or Lua. Interpreters of scripting
languages are available on many different platforms and there is always the
option do define a proprietary scripting language optimized for the problem
domain. Interpreted scripting languages are a effortless and portable way to
make application logic available throughout a \ms{} application. Since any
interpreted code always presents a security issue this aspect has to be examined
with absolute care.

Code Injection using a high level scripting language could also be a convenient
way to introduce \gls{pcg} techniques fluently into the \og{} development
process.

\subsection{Shared Model Automatic Database Integration}

One purpose that the \textit{Shared Model} is designed for is to make it
automatically compatible with the database solution. This is not implemented in
the current version of MicroNet. The proposed design for this feature is to
provide a DataAccess layer via the MicroNet framework which offers a
comprehensive API to query and update objects in the database using the
\textit{Shared Model} as a the driver.

\subsection{Message Listener AST Parsing}

Momentarily MicroNet \textit{Annotations} are quite cumbersome because to fully
describe a message listener all the used message parameters have to be listed.
This aspect could be simplified by parsing the \gls{ast} of the message listener
and to perform reference counting of parameters to automatically deduce
parameters of a message listener.

\subsection{Game Engine Game Model Integration}

The \textit{Shared Model} is additionally designed in a way that is compatible
with the game graph data structure of game engines. It would be possible to
realize an automated integration of the MicroNet \textit{Shared Model} within
the game engine game graph. In a system of this nature game objects which are
currently part of the game simulation could be automatically updated using an
event based messages. The game simulation then could make use of so called views
of game entities and their properties which are automatically updated by the
framework. This way the developer is spared with manual synchronization of the
game state and only needs to contribute code in a framework like fashion to
cover events happening for views. An example uf such an event could be a
\textit{onHealth} event which would cause the game client updates the health bar
visuals accordingly. This dynamic entity view feature would greatly simplify the
networking implementation of the game simulation.
