\chapter{Personal Report}

During this three semester project of which this thesis is the third and final
part I learned a lot about distributed applications. A the beginning of the
first semester I only had basic knowledge about applications that go beyond rich
client applications. Now after two years and three theses I have a belief that
I can design almost any arbitrary cloud application. In this regard i consider
this thesis as a complete success.

One difficulty of this thesis was that it is part of a three semester project
which means that any previous work has a big impact on this thesis. Any mistakes
or critical design decisions will inevitably lead to more work later on which I
could often experience first hand. In other words, completing a three semester
project has proven to be very challenging compared to a one semester project
where the effort is more foreseeable.

The most disappointing matter in this semester was the fact that the focus of
the project shifted more and more away from \og{} related topics in direction of
solely \ms{} related research problems. Since the research problems were often
very general they had hardly anything to do with the game domain in specific.
PCG in this regard had to be neglected almost completely due to time issues. I
would have loved to dive into game specific topics more prominently but leaving
the \ms{} concerns open and half finished was not an option. I therefore decided
to postpone the more game related research to the time after my studies.

But nonetheless I am very pleased with the outcome of this project because I
could complete almost everything that I had planned at the beginning of
this thesis and even the topics that were planned at the start of this project two
years ago could almost all be covered. I especially like how the MicroNet
prototype turned out and I am looking forward to using it in the future to build
awesome \ogs{}. I also plan to continue to work on MicroNet to steadily improve
it.

Finally I owe grateful thanks to my advisor Olaf Zimmermann for his
continuous support during this thesis and the two before this one. He was always
a great support and I could always rely on his assistance when problems
occurred. Thank you for everything, Olaf. 
